---
title: "da460_lab2_grahn"
author: "jason grahn"
date: "10/2/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r data load}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
head(kobe)
```

```{r commands from the handout}
#head of the basket values
kobe$basket[1:9]

#run the calc_streak function to generate the kobe_steak dataset
kobe_streak <- calc_streak(kobe$basket)

#provide a barplot of the kobe_streak data
barplot(table(kobe_streak))

#or do the same tidy style
tibble(kobe_streak) %>% 
  ggplot(aes(x = kobe_streak)) +
  geom_bar()
```
##Exercise 1

#What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0?

A streak length of 1 means that Kobe made one basket and it was immediately followed by a miss; 1 hit and 1 miss. A streak length of zero is a miss that follows another miss.


Notes from the handout: 
"For this lab, we define the length of a shooting streak to be the _number of consecutive baskets made until a miss occurs_."


#Answer these: 
##1. Describe the distribution of streak lengths. What is the typical streak length for this simulated independent shooter with a 45% shooting percentage? How long is the player’s longest streak of baskets in 133 shots?

##2. If you were to run the simulation of the independent shooter a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.

##3. How does Kobe Bryant’s distribution of streak lengths compare to the distribution of streak lengths for the simulated shooter? Using this comparison, do you have evidence that the hot hand model fits Kobe’s shooting patterns? Explain.
```{r}
kobe$basket[1:9]
```

