---
title: "da460_lab2_grahn"
author: "jason grahn"
date: "10/2/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r data load}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
head(kobe)
```

##Exercise 1

###What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0?
```{r commands from the handout}
#head of the basket values
kobe$basket[1:9]
```

A streak length of 1 means that Kobe made one basket and it was immediately followed by a miss; 1 hit and 1 miss. A streak length of zero is a miss that follows another miss.

##Exercise 2

###Describe the distribution of Kobe’s streak lengths from the 2009 NBA finals. What was his typical streak length? How long was his longest streak of baskets?

```{r prep and plot streak data}
#run the calc_streak function to generate the kobe_steak dataset
kobe_streak <- calc_streak(kobe$basket)

#provide a barplot of the kobe_streak data
barplot(table(kobe_streak))

#or do the same tidy style
tibble(kobe_streak) %>% 
  ggplot(aes(x = kobe_streak)) +
  geom_bar()

#but really this is better as a density plot
tibble(kobe_streak) %>% 
  ggplot(aes(x = kobe_streak)) +
  geom_density()
```

I dont think a streak of *zero* should count, so his typical streak length is *1*. His longest streak is 4.

##Exercise 3
###In your simulation of flipping the unfair coin 100 times, how many flips came up heads?

```{r run the sample 100 times}
#build the outcomes vector
outcomes <- c("heads", "tails")

#build a dataset of a fair coin toss
sim_fair_coin <- sample(outcomes, size = 100, replace = TRUE)
table(sim_fair_coin)

#build a dataset of an UNFAIR coin toss
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
table(sim_unfair_coin) 
```

In the unfair coin toss, 26 flips came back heads. 

##Exercise 4
###What change needs to be made to the sample function so that it reflects a shooting percentage of 45%? Make this adjustment, then run a simulation to sample 133 shots. Assign the output of this simulation to a new object called sim_basket.

```{r}
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45, (1-.45)))
table(sim_basket) 
```

```{r comparing kobe against simulated opponent}
table(kobe$basket)
table(sim_basket)
```

