---
title: "da420_lab1_grahn"
author: "jason grahn"
date: "9/23/2018"
output: word_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(psych)
```

##Loading data
```{r load source data}
source('http://www.openintro.org/stat/data/cdc.R') 
```

#Exercise 1: 

##1. how many cases are there in this dataset?
```{r observation count}
nrow(cdc)
```

##2. How many variables in this dataset? 
```{r variable count}
ncol(cdc)
```

##3. For each variable, identify its data type.
```{r}
#but even better than running nrow() and ncol() is glimpse() because it shows both of those AND information about each of the variables
glimpse(cdc)
```

genhlth is categorical 
exerany is categorical (binary)
hlthplan iscategorical (binary)
smoke100 is categorical (binary)
height is continuous and numerical
weight is continuous and numerical
wtdesire is continuous and numerical
age is continuous and numerical
gender is categorical

#Exercise 2

##Create a numerical summary for height and age, and compute the interquartile range for each. 
```{r height summarization}
#the summary of height using base-R
summary(cdc$height)
```
The IQR is `r (70 - 64) * 1.5`.

```{r age summarization}
#the summary of age using base-R
summary(cdc$age)
```
the IQR is `r (57 - 31) * 1.5`.

```{r doing this the easy way}
#or if we want to gather summary stats for both at once, we can use tidyverse dplyr select with psych::describe() 
cdc %>% 
  select(height, age) %>% 
  describe(IQR = TRUE)
```

Compute the relative frequency distribution for gender and exerany. 
```{r}
table(cdc$gender)/nrow(cdc)
```

How many males are in the sample? 
```{r HOW MANY MALES}
cdc %>%  
  select(gender) %>% 
  filter(gender == "m") %>% 
  summarize(count_of_males = n())

#or
table(cdc$gender)
#which answers for both genders
```

What proportion of the sample reports being in *excellent* health?
```{r}
cdc %>% 
  select(genhlth) %>% 
  group_by(genhlth) %>% 
  summarise(health_count = n()) %>%
  mutate(freq = health_count / nrow(cdc) ) %>% 
  filter(genhlth == "excellent")

#which is maybe a little bit long winded, so in base-R we can create the same using a couple of commands
table(cdc$genhlth)/nrow(cdc) 
```


