---
title: "lab5"
author: "jason grahn"
date: "10/24/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
library(psych)

download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")

```

#R component

```{r}
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)

population.tidy <- ames %>% 
  select(Gr.Liv.Area)

samp.tidy <- population.tidy %>% 
  rep_sample_n(size = 60, reps = 1)

summary(samp.tidy)
```

##Exercise 1
###Describe the distribution of your sample. What would you say is the “typical” size within your sample? Also state precisely what you interpreted “typical” to mean.
```{r}
samp.describe <- describe(samp.tidy$Gr.Liv.Area)

samp.histogram <- samp.tidy %>% 
  ggplot() +
  geom_histogram(aes(x = Gr.Liv.Area), bins = 15)

samp.qqplot <- samp.tidy %>% 
  ggplot(aes(sample = Gr.Liv.Area)) + 
  stat_qq() +
  stat_qq_line(color = "red")

Mode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
     }

Mode( samp.tidy$Gr.Liv.Area)

samp.describe
samp.histogram
samp.qqplot
```

The sample appears somewhat normal but with a very strong right skew as we can tell with a skewness of `1.02`. The `median` is 1498 with a `mean` of 1574. 

##Exercise 2
###Would you expect another student’s distribution to be identical to yours? Would you expect it to be similar? Why or why not?

I definitely would _NOT_ expect another students distribution to be identical because we pulled the descriptive statistics from a sample of 60 out of a population of 2,930. If the population is truly normal, then their sample would ideed be similar. 

##Exercise 3
###For the confidence interval to be valid, the sample mean must be normally distributed and have standard error s/n‾√. What conditions must be met for this to be true?

For the confidence interval of a sample to be true, the sample must be random and independent. The sample size needs to be greater than 30, and there should be no skewness to the distribution.

##Exercise 4
###What does “95% confidence” mean? If you’re not sure, see Section 4.2.2.

The term "95% confidence" means that 95% of the time our point estimates will be within 2 standard errors of the parameter for that statistic.

##Exercise 5
###Does your confidence interval capture the true average size of houses in Ames? If you are working on this lab in a classroom, does your neighbor’s interval capture this value?

```{r}
pop.mean <- mean(population.tidy$Gr.Liv.Area) 

#baseR
sample_mean <- mean(samp)
se <- sd(samp) / sqrt(60)
lower <- sample_mean - 1.96 * se
upper <- sample_mean + 1.96 * se
c(lower, upper)
```

The population mean is `r pop.mean`. The confidence interval of my sample captures the population mean, yes. I am not working on this within a classroom so I am unable to answer for my neighbors.


##Exercise 6
###Each student in your class should have gotten a slightly different confidence interval. What proportion of those intervals would you expect to capture the true population mean? Why? If you are working in this lab in a classroom, collect data on the intervals created by other students in the class and calculate the proportion of intervals that capture the true population mean.

We aren't in a classroom. I would expect 95% of the intervals to capture the true population mean because that's how confidence intervals work. 

update:

ok so i want to figure out how I can do this for the entire class

```{r}
#load packages
library(tidyverse)
library(broom)
library(psych)

#download and load file
download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")

#get the population into a tidy data frame
population.tidy <- ames %>% 
  select(Gr.Liv.Area)

#what's the population mean?
population.mean <- mean(population.tidy$Gr.Liv.Area)

#run the sampling of 60 for 60 students
class.samp.tidy <- population.tidy %>% 
  rep_sample_n(size = 60, reps = 60)

#whats the mean for all the samples taken by the simulated 60 students?
class.mean <- class.samp.tidy %>% 
  #group by the replicate column
  group_by(replicate) %>% 
  #give me an average per replicate
  summarise(sample_mean = mean(Gr.Liv.Area)) %>% 
  ungroup() %>% 
  #give me the average for all averages
  summarise(sample_mean = mean(sample_mean))

#ok, lets make a pretty picture with all of this...
class.samp.tidy %>% 
  group_by(replicate) %>% 
  #build our confidence intervals
  summarise(sample_mean = mean(Gr.Liv.Area),
            standard_dev = SD(Gr.Liv.Area),
            se = standard_dev/sqrt(60),
            lower = sample_mean - 1.96 * se,
            upper = sample_mean + 1.96 * se) %>% 
  #and now for the pretty picture
  ggplot() +
  theme_minimal() +
  geom_pointrange(aes(x = factor(replicate), y = sample_mean, ymin = lower, ymax = upper), alpha = 0.5) + 
  geom_hline(yintercept = population.mean, color = "red") +
  geom_hline(yintercept = class.mean$sample_mean, color = "blue") +
  coord_flip() +
  labs(title = "Confidence intervals for 60 simulated students",
       subtitle = "Red line is population mean, blue line is mean of sample means")
```

nice.

```{r}
DescTools::MeanCI(class.samp.tidy$Gr.Liv.Area,
                  conf.level = 0.95,
                  sides = "two.sided")

library(Rmisc)
by.replicate <- class.samp.tidy %>% 
  group_by(replicate) 

by.replicate %>% 
  do(CI(, ci = 0.95))

library(broom)
tibble(a = list(c(1, 2), c(3, 4))) %>%
  mutate(ttest = map(a, t.test, conf.level = 0.95) %>% 
           map(tidy) %>% 
           map(~ select(., conf.low, conf.high))) %>%
  unnest(ttest, .drop = FALSE)

class.samp.tidy %>% 
  group_by(replicate) %>% 
  nest(Gr.Liv.Area) %>%
  mutate(ttest = map(replicate, t.test, conf.level = 0.95) %>% 
           map(tidy) %>% 
           map(~ select(., conf.low, conf.high))) %>%
  unnest(ttest, .drop = FALSE)

```

