---
title: "lab5"
author: "jason grahn"
date: "10/24/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
library(psych)

download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")

```

#R component

```{r}
population <- ames$Gr.Liv.Area
samp <- sample(population, 60)

population.tidy <- ames %>% 
  select(Gr.Liv.Area)

samp.tidy <- population.tidy %>% 
  rep_sample_n(size = 60, reps = 1)
```

##Exercise 1
###Describe the distribution of your sample. What would you say is the “typical” size within your sample? Also state precisely what you interpreted “typical” to mean.
```{r}
samp.describe <- describe(samp.tidy$Gr.Liv.Area)

samp.histogram <- samp.tidy %>% 
  ggplot() +
  geom_histogram(aes(x = Gr.Liv.Area), bins = 15)

samp.qqplot <- samp.tidy %>% 
  ggplot(aes(sample = Gr.Liv.Area)) + 
  stat_qq() +
  stat_qq_line(color = "red")

samp.describe
samp.histogram
samp.qqplot
```

The sample appears somewhat normal but with a very strong right skew as we can tell with a skewness of `1.02`. The `median` is 1498 with a `mean` of 1574. 

##Exercise 2
###Would you expect another student’s distribution to be identical to yours? Would you expect it to be similar? Why or why not?

I definitely would _NOT_ expect another students distribution to be identical because we pulled the descriptive statistics from a sample of 60 out of a population of 2,930. If the population is truly normal, then their sample would ideed be similar. 

##Exercise 3
###For the confidence interval to be valid, the sample mean must be normally distributed and have standard error s/n‾√. What conditions must be met for this to be true?

For the confidence interval of a sample to be true, the sample must be random and independent. The sample size needs to be greater than 30, and there should be no skewness to the distribution.

##Exercise 4
###What does “95% confidence” mean? If you’re not sure, see Section 4.2.2.

The term "95% confidence" means that 95% of the time our point estimates will be within 2 standard errors of the parameter for that statistic.

##Exercise 5
###Does your confidence interval capture the true average size of houses in Ames? If you are working on this lab in a classroom, does your neighbor’s interval capture this value?

```{r}
pop.mean <- mean(population.tidy$Gr.Liv.Area) 

#baseR
sample_mean <- mean(samp)
se <- sd(samp) / sqrt(60)
lower <- sample_mean - 1.96 * se
upper <- sample_mean + 1.96 * se
c(lower, upper)
```

The population mean is `r pop.mean`. The confidence interval of my sample captures the population mean, yes. I am not working on this within a classroom so I am unable to answer for my neighbors.


##Exercise 6
###Each student in your class should have gotten a slightly different confidence interval. What proportion of those intervals would you expect to capture the true population mean? Why? If you are working in this lab in a classroom, collect data on the intervals created by other students in the class and calculate the proportion of intervals that capture the true population mean.

We aren't in a classroom. I would expect 95% of the intervals to capture the true population mean because that's how confidence intervals work. 

