---
title: "midterm"
author: "jason grahn"
date: "10/26/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
library(psych)
```

#Problem 1:
##N. 1 
Download smoking.csv (or smoking.txt) and read corresponding data into R. Example command in R: `MyData <- read.csv(file="path/TheDataIWantToReadIn.csv",header=TRUE, sep=",")`Make sure to include the code/command.

```{r loading smoking data}
library(readr)
smoking <- read_csv("smoking.csv")
#View(smoking)
```

##N. 2 
How many observations are there in this data set? 
How many variables, and what are they? 

```{r}
#how many observations and variables are shown with glimpse()
glimpse(smoking)
```

The `glimpse()` function shows us both the number of variables and rows. The number of rows in the data is `r nrow(smoking)`, and the number of variables is 12. They are:
    
    * gender
    * age
    * maritalStatus
    * highestQualification
    * nationality
    * ethnicity
    * grossIncome
    * region
    * smoke
    * amtWeekends
    * amtWeekdays
    * type

```{r}
#What is the 300th observation of nationality?
smoking$nationality[[300]]
```

##N. 3 
Create a numerical summary for age and compute the interquartile range. 
```{r}
summary(smoking$age)

age.iqr <- describe(smoking$age, IQR = TRUE) %>% 
  select(IQR)
```

The numerical summary for `age` is shown above. The IQR is `r age.iqr$IQR`.

Compute the relative frequency distribution for gender. 
```{r}
table(smoking$gender)/nrow(smoking)
```

How many males are in the sample? Include both the code/command and the output/graph.
```{r}
male.count <- nrow(
  smoking %>% 
    filter(gender == "Male")
)
```

There are `r male.count` males in the sample. 

##N. 4
Using numerical summaries and a side-by-side box plot, determine if male smokers are as old as female smokers. Include both the code/command and the output/graph.

```{r setup smoking filtered data}
s.g <- smoking %>% 
  filter(smoke == "Yes") %>% 
  select(age, gender)

#group the male data together
s.g.m <- s.g %>% filter(gender == "Male")

#group the female data together
s.g.f <- s.g %>% filter(gender == "Female")
```

```{r summary of make smoker ages}
#age summary for male smokers
summary(s.g.m$age)
```

```{r summary of female smoker ages}
#age summary for female smokers
summary(s.g.f$age)
```

```{r side by side boxplot of smoker ages}
s.g %>% 
  ggplot(aes(x = gender, y = age)) + 
  geom_boxplot() +
  geom_jitter(alpha = 0.1, width = 0.05)
```

##N. 5
Create a bar chart or frequency table for maritalStatus, what is the proportion for Divorced, Single, Married, and Widowed, respectively? What can you interpret from these numbers? Include both the code/command and the output/graph.

```{r status frequency table}
table(smoking$maritalStatus)/nrow(smoking) 
```

We can interpret from these numbers that most people in the smokers data frame are married. Single is next, then widowed. 

#Problem 2

#N. 1
Suppose we’re flipping an unfair coin that we know only lands heads 30% of the time. Please simulate this flip 10 times, what is the proportion of heads? 
```{r}

```

If you simulate this flip 100 times, what is the proportion of heads now? Include both the code/command and the output/graph.