---
title: "lab5"
author: "jason grahn"
date: "10/24/2018"
output: word_document
---

#Inference for numerical data
```{r load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(moderndive)
library(psych)
```

```{r load data}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
glimpse(nc)
```


##1
###What are the cases in this data set? How many cases are there in our sample?

The cases in this dataset, otherwise known as observations, are births recorded in North Carolina. `glimpse()` shows us there are 1000 observations. We can also run summary() and review any of the categorical variables like `mature`. The summation of which comes to 1000.
```{r}
summary(nc)
```

##2
###Make a side-by-side boxplot of habit and weight. What does the plot highlight about the relationship between these two variables?
```{r}
nc %>% 
  #There are NAs in the data that need to be removed.
  filter(!is.na(habit)) %>% 
  #now plot the results.
  ggplot(aes(x = habit, y = weight)) +
  geom_boxplot() +
  theme_light()

by(nc$weight, nc$habit, mean)
```

We definitely have an observed difference. While the medians are similar, they *are* different. The boxplot shows smokers have a tighter tolerance for weight, and that non-smokers have much more variation with weights.

##3
###Check if the conditions necessary for inference are satisfied. Note that you will need to obtain sample sizes to check the conditions. You can compute the group size using the same by command above but replacing mean with length.
```{r}
by(nc$weight, nc$habit, length)
```
 
The conditions seem to be met: The 1,000 observations are likely less than 10% of the population to ensure the sampling is random, which fulfills independence. We have more than 30 samples, which satifies large number central limit theorem. Though distributions shown on box and whisker plots have outliers, these seem reasonable for the sample size. 

##4
###Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.

    * H0 = The average weights of babies born to smoking and non-smoking mothers are not different
    * HA = The average weights of babies born to smoking and non-smoking mothers are different
    
##5
###Change the type argument to “ci” to construct and record a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers.
```{r}
#inference(y = nc$weight, 
#          x = nc$habit, 
#          est = "mean", 
#          type = "ht", 
#          null = 0, 
#          alternative = "twosided", 
#          method = "theoretical")

inference(y = nc$weight, 
          x = nc$habit, 
          est = "mean", 
          type = "ci", 
          null = 0, 
          alternative = "twosided", 
          method = "theoretical")
```
The `95% Confidence interval` is `(0.0534 , 0.5777 )` and this command also ends up making the same boxplot we made before. 

#Inference for categorical data

##1 
###In the first paragraph, several findings are reported. Do these percentages appear to be _sample statistics_ (derived from the data sample) or _population parameters_?

The _second_ paragraph states: 
    
    * 59% of the world said that they think of themselves as religious person, 23% think of themselves as not religious whereas 13% think of themselves as convinced atheists. However, when we compare this to the Irish population, only 47% consider themselves religious, placing Ireland low on the index of being religious coming in at position 43 out of 57 countries. 
    
It is not feasible to conduct this level of analysis with _every_ individual in the world. These appear to be sample statistics. 

##2
###The title of the report is “Global Index of Religiosity and Atheism”. To generalize the report’s findings to the global human population, what must we assume about the sampling method? Does that seem like a reasonable assumption?

We must assume the this is a random sample that has a high enough _n_ to account for a reasonable percentage of the global population and that each interviewee / respondant was part of a simple random selection process. 

##3
###What does each row of Table 6 correspond to? What does each row of atheism correspond to?
```{r download part 2 data} 
download.file("http://www.openintro.org/stat/data/atheism.RData", destfile = "atheism.RData")
load("atheism.RData")
glimpse(atheism)
#View(atheism)
```

The rows in `Table 6` of [the article](https://sidmennt.is/wp-content/uploads/Gallup-International-um-tr%C3%BA-og-tr%C3%BAleysi-2012.pdf) is titled `GLOBAL RELIGIOSITY AND ATHEISM INDEX FOR 2012`. The table provides a country, the sample size of those surveyed, and the percent of respondants as they self-selected their level of religiousity. 

The observations in the `atheism` data are a given country, the year of the survey, and the respondants selection of atheist or non-atheist. This appears to be individual line-item responses.
```{r}
nat.summary <- atheism %>% 
       group_by(nationality, response) %>% 
       summarize(count = n())

head(nat.summary, 10)
```

##4
###Using the command below, create a new dataframe called us12 that contains only the rows in atheism associated with respondents to the 2012 survey from the United States. 
```{r}
us12 <- subset(atheism, nationality == "United States" & year == "2012")

#or the tidy version of same
us12.tidy <- atheism %>% 
  filter(nationality == "United States",
         year == "2012")
```

###Next, calculate the proportion of atheist responses. 
```{r}
us12atheists <- us12.tidy %>% 
  group_by(response) %>% 
  summarise(response_count = n()) %>%
  mutate(proportion = response_count / nrow(us12.tidy)) %>% 
  filter(response == "atheist")

us12atheists
```

###Does it agree with the percentage in Table 6? If not, why?
Yes, this agrees with the percentage in Table 6, which shows a US atheist population of `5%`. The proportion of atheists by calculation of the openintro data is `4.99%`, which is close enough to 5% to make this acceptable as agreeing.  

##5
###Write out the conditions for inference to construct a 95% confidence interval for the proportion of atheists in the United States in 2012. Are you confident all conditions are met?
The conditions are:

    1. The observations are independent. This condition is met because assume that the samples are randomly selected. 
    2. This also requires a sample size of np >= 10 and n(1-p)>=10. This condition is met.
        * np = 1002*0.05 = 50, which is >= 10. 
        * n(1-p) = 1002*0.95 = 952, which is also >= 10.
        
```{r}
us12$response %>% 
inference(.,
          est = "proportion", 
          type = "ci", 
          method = "theoretical", 
          success = "atheist")

#describe(us12$response)
```

##6
###Based on the R output, what is the margin of error for the estimate of the proportion of the proportion of atheists in US in 2012?

Assuming a 95% confidence interval: 
```{r}
#we'll need to make the confidence interval value for .95
a <- 1 - 0.05/2
ci.95 <- qnorm(a)

#one way is to use the SE from above and manufacture the Margin of error
margin.error <- ci.95* 0.0069
margin.error

#but that's not the best for reproducibility. Let's make the code do the thinking
p <- us12atheists$proportion
q <- 1-p                                    

#the use P*Q to make the standard error
stderr <- sqrt( (p*q) / nrow(us12.tidy) )
#stderr

#then use the confidence interval to make the margin of error:
margerr <- ci.95 * stderr 
margerr
```

The margin of error is either `r margin.error` or `r margerr` depending on how it's calculated. Either way we can round this to 0.0135.

##7
###Using the inference function, calculate confidence intervals for the proportion of atheists in 2012 in two other countries of your choice, and report the associated margins of error. Be sure to note whether the conditions for inference are met. It may be helpful to create new data sets for each of the two countries first, and then use these data sets in the inference function to construct the confidence intervals.
```{r}

```


